Designing Types for R, Empirically

### 2020-09-14 paper/main.tex
new words in. let the edits commence

\begin{document}
\title{Designing Types for R, Empirically}
\author{Alexi Turcotte}\affiliation{\institution{Northeastern University}}
\author{Aviral Goel}\affiliation{\institution{Northeastern University}}
\author{Filip Krikava}\affiliation{\institution{Czech Technical University in Prague}}
\author{Jan Vitek}\affiliation{\institution{Nortjeastern / Czech Technical University}}

\newcommand{\NUMFUNCTIONS}{25,215\xspace}  %%% TODO auto-gen
\newcommand{\NUMPACKAGES}{412\xspace}  %%% TODO auto-gen
\newcommand{\PACKAGES}{412\xspace}  %%% TODO auto-gen
\newcommand{\genthat}{genthat\xspace}  %%% TODO auto-gen
\newcommand{\YEARS}{20\xspace} %%TODO fix
\newcommand{\PERCFAILEDASSERTIONS}{2.05\%\xspace}
\newcommand{\PERCFAILEDASSERTIONSWUNDEF}{1.98\%\xspace}
\newcommand{\PERCASSERTIONSUNDEF}{0.07\%\xspace}
\newcommand{\NUMPKGSEVAL}{8,694\xspace} % autogenerate
\newcommand{\NUMASSERTIONS}{98,105,161\xspace} %autogenerate
\newcommand{\PROPFUNSFAILEDCHECK}{12.29\%\xspace}
\newcommand{\PROPFUNSFAILEDCHECKNOSTHREE}{10.30\%\xspace} 
\newcommand{\PROPARGSFAILINGASSERTS}{2.40\%\xspace} % autogen....
\newcommand{\PERCSUCCARG}{97.60\%\xspace}
\newcommand{\PERCSUCCFUNNOSTHREE}{89.70\%\xspace}
\newcommand{\PERCSUCCFUNS}{87.70\%\xspace}
\newcommand{\PERCCALLSBADFUNSINTESTS}{2.73\%\xspace}
\newcommand{\TOTNUMSIGS}{21,730\xspace}
\newcommand{\PROPSTRUCTSDYN}{65.97\%\xspace}
\newcommand{\NUMFILESRUN}{160,379\xspace}
\newcommand{\NUMFILESREVDEPS}{153,806\xspace}
\newcommand{\NUMFILESKAGGLE}{6,573\xspace}

@@

\AT{TODO: Softer intro to R.
People always get confused about c() and stuff like that.
We have some space, we can take it.}

@@

@@

\begin{table}
% BEGIN Autogenerated
\small
\begin{tabular}{llrrr}
\toprule
Passed & Arg Type & Occurrences & \% Total & Cumulative \%\\
\midrule
\VEC\dbl & \CLASS{\texttt{matrix}} & 705,036 & 36.36 & 36.36 \\ 
\VEC\dbl & \VEC\int & 189,800 & 9.79 & 46.15 \\ 
\chr & \CLASS{\texttt{bignum}} | \VEC\raw & 100,100 & 5.16 & 51.31 \\
\CLASS{{\tt simpleError, error, condition}} & \CLASS{{\tt data.frame}} | \CLASS{{\tt matrix}} | \CLASS{{\tt randomForest}} | \VEC\dbl & 78,197 & 4.03 & 55.34 \\ 
  \CLASS{{\tt data.frame}} & \CLASS{{\tt data.frame, tbl, tbl\_df}} & 58,809 & 3.03 & 58.37 \\ 
  \CLASS{{\tt matrix}} & \CLASS{{\tt timeSeries}} & 53,482 & 2.76 & 61.13 \\ 
 \VEC\dbl & \VEC\int & 33,350 & 1.72 & 62.85 \\ 
 \dbl & \CLASS{{\tt data.frame}} & 32261 & 1.66 & 64.52 \\ 
  \VEC\dbl & \CLASS{{\tt data.frame}} & 31361 & 1.62 & 66.13 \\ 
  \CLASS{{\tt xml\_node}} & \CLASS{{\tt xml\_missing}} $|$ \CLASS{{\tt xml\_nodeset}} $|$ \CLASS{{\tt xml\_document, xml\_node}} & 30,330 & 1.56 & 67.70 \\ 
\bottomrule
\end{tabular}
% END Autogenerated
\caption{Top contract failures}
\label{tbl:fail-breakdown-0}
\end{table}

@@

 \begin{table}
% BEGIN Autogenerated
\small
\begin{tabular}{llrrr}
\toprule
Passed & Arg Type & \# Args Failed & \# Args with Type & \% Failure\\
\VEC\dbl & \CLASS{\texttt{matrix}} & 18 & 1522 & 1.18 \\ 
\VEC\dbl & \VEC\int  & 66 & 5865 & 1.13 \\ 
\chr & \CLASS{\texttt{bignum}} | \VEC\raw & 1 & 3 & 33.33 \\ 
\CLASS{{\tt simpleError, error, condition}} & \CLASS{{\tt data.frame}} | \CLASS{{\tt matrix}} | \CLASS{{\tt randomForest}} | \VEC\dbl & 2 & 2 & 100 \\ 
  \CLASS{{\tt data.frame}} & \CLASS{{\tt data.frame, tbl, tbl\_df}} & 23 & 196 & 11.73 \\ 
  \CLASS{{\tt matrix}} & \CLASS{{\tt timeSeries}} & 21 & 39 & 53.85 \\ 
 \VEC\dbl & \VEC\int & 31 & 624 & 4.97 \\ 
 \dbl & \CLASS{{\tt data.frame}} & 2 & 1025 & 0.20 \\ 
  \VEC\dbl & \CLASS{{\tt data.frame}}  & 6 & 1025 & 0.59 \\ 
  \CLASS{{\tt xml\_node}} & \CLASS{{\tt xml\_missing}} $|$ \CLASS{{\tt xml\_nodeset}} $|$ \CLASS{{\tt xml\_document, xml\_node}}  & 1 & 1 & 100 \\ 
\bottomrule
\end{tabular}

@@

 \begin{table}
% BEGIN Autogenerated\
\small
\begin{tabular}{llrrr}
\toprule
Passed & Arg Type & \# Args Failed & \# Args with Type & \% Failure\\
\midrule
\VEC\dbl & \VEC\int & 31 & 624 & 4.97 \\
\dbl & \int & 21 & 519 & 4.05\\
\VEC\chr & \chr | \null & 10 & 256 & 3.91 \\
\NAVEC\lgl & \VEC\lgl & 8  & 219 & 3.65 \\
\NASCA\lgl & \VEC\lgl & 5 & 219 & 2.28 \\
\bottomrule
\end{tabular}
% END Autogenerated
 \caption{What is the highest failure rate among popular argument types? For argument signatures whose frequency is in the 90th percentile}
 \label{tbl:fail-breakdown-2}
 \end{table}

  \begin{table}
% BEGIN Autogenerated
\small
\begin{tabular}{ll@{}rrr@{}}
\toprule
Passed & Arg Type & \#Args Fail & \#Args w Type & \% Failure\\
\midrule
\CLASS{{\tt matrix}} & \CLASS{{\tt timeSeries}} & 21 & 39 & 35.90 \\
\VEC\dbl & \CLASS{{\tt timeSeries}} & 21 & 39 & 35.90 \\
\CLASS{{\tt data.frame}} & \CLASS{{\tt timeSeries}} & 14 & 39 & 35.90 \\
\CLASS{{\tt dtplyr\_step\_first, dtplyr\_step}} & \CLASS{{\tt data.frame}} | Grouped Tibble | Tibble & 10 & 45 & 22.22 \\
\chr & \raw & 6 & 31 & 19.35 \\
% \CLASS{\texttt{data.frame}} & \CLASS{\texttt{data.frame}, \texttt{tbl}, \texttt{tbl\_df}} & 63 & 284 & 22.18\\
% \CLASS{\texttt{matrix}} & \CLASS{\texttt{array}} & 13 & 70 & 18.57\\
% \NAVEC\lgl & \lgl | \null & 9 & 49 & 18.37\\
% \VEC\chr & \LIST{\VEC\chr} & 20 & 109 & 18.35\\
% \VEC\clx & \clx & 7 & 42 & 16.67\\
\bottomrule
\end{tabular}
% END Autogenerated
 \caption{What is the highest failure rate among popular argument types? For argument signatures whose frequency is in the 80th percentile.
 *: Grouped Tibble is \CLASS{{\tt data.frame, grouped\_df, tbl, tbl\_df}} , and Tibble is \CLASS{{\tt data.frame, grouped\_df, tbl, tbl\_df}}.
 \AT{TODO Fix this table.}}
 \label{tbl:fail-breakdown-3}
 \end{table}

@@

 First, we were curious about how many \null-typed values flowed into non-nullable arguments.
 We found that \AT{fill in numbers}.
 We manually inspected some of the offending functions and observed three main patterns.

 First, we found that many of these errors occurred in arguments that have a {\tt NULL} default value.
 This would be the case when the programmer only tested the function by passing values to these null-by-default arguments, and clients of the package make use of the default.
 As an example, we observed a contract assertion failure when {\tt inner} and {\tt labels} were {\tt NULL} in the following function:
 \begin{lstlisting}
nlme::nfGroupedData <- function (formula, data = NULL, order.groups = TRUE, 
    FUN = function(x) max(x, na.rm = TRUE), outer = NULL, inner = NULL, 
    labels = NULL, units = NULL) {
  # ...
}
 \end{lstlisting}
 In fact, we observed \AT{something about coverage.}

@@

 Next, we were curious about how often vectors were passed to arguments expecting scalars.
 We found \AT{finish analysis, numbers}.
 Besides an outright error, this kind of contract assertion failure might indicate that a function was not well-tested, in that it was only ever tested with unit-length vectors being passed to an argument which is intended to have a vector type.
 Further, these errors may reveal functions that were not designed with a vector-typed argument in mind, but can in fact handle vectors of values (in R, most functions that can accept scalars can also accept vectors).
 As an example, consider the function \code{BBmisc::strrepeat}, which takes a string and repeats it a specified number of times:
 \begin{lstlisting}
BBmisc::strrepeat <- function (x, n, sep = "") {
    paste0(rep.int(x, n), collapse = sep)
}

@@

Finally, we were curious to see what patterns of errors occurred in arguments expecting classes.
\AT{Overall, the numbers are ...}.
Many of these errors appear to be related to ``subclasses'' being passed to arguments expecting a ``superclass''.
For instance, there are \AT{many} instances of \CLASS{{\tt data.frame}} being passed to arguments expecting tibbles or data.tables (which can be readily converted).
\AT{Check dispatch, and maybe count it out when doing this.}

Other common patterns include \CLASS{{\tt matrix}}, \CLASS{{\tt data.frame}}, and vectors being passed to arguments expecting \CLASS{{\tt array}}, which is \AT{elaborate on arrays}.
\AT{Fill rest of this in: vectors and data frames passed to matrices. reiterate how this coercion is ad hoc. maybe talk about coercion semantics.}

@@

 \AT{TODO: Fill in and discuss Kaggle results.}

 To further validate our inferred types, we repeated the experiment discussed in this section on Kaggle code.
 By-and-large, we saw no meaningful difference in the two data sets, with the contract assertion failure patterns being repeated from the reverse dependencies.

 \AT{Discuss how things match up between Kaggle and revdeps.}

\begin{itemize}
  \item compare which functions have failures 
  \item look at the coverage 
  \item types of errors?
\end{itemize}

 %
 %
 \subsubsection{Working Failed Assertions into Types}

 \AT{TODO: compare with types that we get when looking at reverse dependencies, and discuss implications.}

Throughout this section, we have seen instances where the reverse dependencies of a package exercise its functions in a way that was not accounted for in the package's own test and  example code.
These do not always represent errors, and may represent cases of oversight by package writers.

If we {\it include} running a package's reverse dependencies in our type inference process, we will obtain different types for certain functions.
In doing so, we observe \ldots

@@

\AT{Talk about how we thought do to a different unification for class types.}

>>>>

\begin{document}
\title{Designing Types for R, Empirically}
\author{Alexi Turcotte}\affiliation{\institution{Northeastern University}}
\author{Aviral Goel}\affiliation{\institution{Northeastern University}}
\author{Filip Krikava}\affiliation{\institution{Czech Technical University in Prague}}
\author{Jan Vitek}\affiliation{\institution{Northeastern / Czech Technical University}}

\newcommand{\NUMFUNCTIONS}{25,215\xspace}  %%% TODO auto-gen
\newcommand{\NUMPACKAGES}{412\xspace}  %%% TODO auto-gen
\newcommand{\PACKAGES}{412\xspace}  %%% TODO auto-gen
\newcommand{\genthat}{genthat\xspace}  %%% TODO auto-gen
\newcommand{\YEARS}{20\xspace} %%TODO fix
\newcommand{\PERCFAILEDASSERTIONS}{2.05\%\xspace}
\newcommand{\PERCFAILEDASSERTIONSWUNDEF}{1.98\%\xspace}
\newcommand{\PERCASSERTIONSUNDEF}{0.07\%\xspace}
\newcommand{\NUMPKGSEVAL}{8,694\xspace} % autogenerate
\newcommand{\NUMASSERTIONS}{98,105,161\xspace} %autogenerate
\newcommand{\PROPFUNSFAILEDCHECK}{12.29\%\xspace}
\newcommand{\PROPFUNSFAILEDCHECKNOSTHREE}{10.30\%\xspace} 
\newcommand{\PROPARGSFAILINGASSERTS}{2.40\%\xspace} % autogen....
\newcommand{\PERCSUCCARG}{97.60\%\xspace}
\newcommand{\PERCSUCCFUNNOSTHREE}{89.70\%\xspace}
\newcommand{\PERCSUCCFUNS}{87.70\%\xspace}
\newcommand{\PERCCALLSBADFUNSINTESTS}{2.73\%\xspace}
\newcommand{\TOTNUMSIGS}{21,730\xspace}
\newcommand{\PROPSTRUCTSDYN}{65.97\%\xspace}
\newcommand{\NUMFILESRUN}{160,379\xspace}
\newcommand{\NUMFILESREVDEPS}{153,806\xspace}
\newcommand{\NUMFILESKAGGLE}{6,573\xspace}
\newcommand{\NUMNULLASSERTERRORS}{6.46\%\xspace}
\newcommand{\NUMFUNSWITHNULLASSERTERROR}{141\xspace}
\newcommand{\PERCDYNAMICVECERRORS}{12.73\%\xspace}
\newcommand{\NUMFUNSWITHVECASSERTERROR}{114\xspace}
\newcommand{\PERCFAILASSCLASS}{81.44\%\xspace}
\newcommand{\PERCFAILEDONECLASS}{51.13\%\xspace}

% Kaggle
 \newcommand{\PERCFAILEDASSERTIONSKAGGLE}{2.14\%\xspace}
\newcommand{\PERCFAILEDASSERTIONSKAGGLENOUNDEF}{0.42\%\xspace}
 \newcommand{\PROPFUNSWITHFAILEDASSERTKAGGLE}{15.98\%\xspace}
 \newcommand{\NUMTOTALASSERTIONSKAGGLE}{1,9038,496\xspace}
 \newcommand{\NUMASSERTEDFUNCTIONSKAGGLE}{970\xspace}

@@

@@

\AT{TODO: Update plots, reread and readjust wording and examples since some tables have changed.}

@@

\begin{table}
% BEGIN Autogenerated
\scriptsize
\begin{tabular}{llrrr}
\toprule
Passed & Arg Type & Occurrences & \% Total & Cumulative \%\\
\midrule
\VEC\dbl & \CLASS{\texttt{matrix}} & 705,036 & 36.36 & 36.36 \\ 
\VEC\dbl & \VEC\int & 189,800 & 9.79 & 46.15 \\ 
\chr & \CLASS{\texttt{bignum}} | \VEC\raw & 100,100 & 5.16 & 51.31 \\
\CLASS{{\tt simpleError, error, condition}} & \CLASS{{\tt data.frame}} | \CLASS{{\tt matrix}} | \CLASS{{\tt randomForest}} | \VEC\dbl & 78,197 & 4.03 & 55.34 \\ 
  \CLASS{{\tt data.frame}} & \CLASS{{\tt data.frame, tbl, tbl\_df}} & 58,809 & 3.03 & 58.37 \\ 
  \CLASS{{\tt matrix}} & \CLASS{{\tt timeSeries}} & 53,482 & 2.76 & 61.13 \\ 
 \VEC\dbl & \VEC\int & 33,350 & 1.72 & 62.85 \\ 
 \dbl & \CLASS{{\tt data.frame}} & 32261 & 1.66 & 64.52 \\ 
  \VEC\dbl & \CLASS{{\tt data.frame}} & 31361 & 1.62 & 66.13 \\ 
  \CLASS{{\tt xml\_node}} & \CLASS{{\tt xml\_missing}} $|$ \CLASS{{\tt xml\_nodeset}} $|$ \CLASS{{\tt xml\_document, xml\_node}} & 30,330 & 1.56 & 67.70 \\ 
\bottomrule
\end{tabular}
% END Autogenerated
\caption{Top contract failures}
\label{tbl:fail-breakdown-0}
\end{table}

@@

 \begin{table}
% BEGIN Autogenerated
\scriptsize
\begin{tabular}{llrrr}
\toprule
Passed & Arg Type & \# Args Failed & \# Args with Type & \% Failure\\
\VEC\dbl & \CLASS{\texttt{matrix}} & 18 & 1522 & 1.18 \\ 
\VEC\dbl & \VEC\int  & 66 & 5865 & 1.13 \\ 
\chr & \CLASS{\texttt{bignum}} | \VEC\raw & 1 & 3 & 33.33 \\ 
\CLASS{{\tt simpleError, error, condition}} & \CLASS{{\tt data.frame}} | \CLASS{{\tt matrix}} | \CLASS{{\tt randomForest}} | \VEC\dbl & 2 & 2 & 100 \\ 
  \CLASS{{\tt data.frame}} & \CLASS{{\tt data.frame, tbl, tbl\_df}} & 23 & 196 & 11.73 \\ 
  \CLASS{{\tt matrix}} & \CLASS{{\tt timeSeries}} & 21 & 39 & 53.85 \\ 
 \VEC\dbl & \VEC\int & 31 & 624 & 4.97 \\ 
 \dbl & \CLASS{{\tt data.frame}} & 2 & 1025 & 0.20 \\ 
  \VEC\dbl & \CLASS{{\tt data.frame}}  & 6 & 1025 & 0.59 \\ 
  \CLASS{{\tt xml\_node}} & \CLASS{{\tt xml\_missing}} $|$ \CLASS{{\tt xml\_nodeset}} $|$ \CLASS{{\tt xml\_document, xml\_node}}  & 1 & 1 & 100 \\ 
\bottomrule
\end{tabular}

@@

 \begin{table}
% BEGIN Autogenerated\
\scriptsize
\begin{tabular}{llrrr}
\toprule
Passed & Arg Type & \# Args Failed & \# Args with Type & \% Failure\\
\midrule
\VEC\dbl & \VEC\int & 31 & 624 & 4.97 \\
\dbl & \int & 21 & 519 & 4.05\\
\VEC\chr & \chr | \null & 10 & 256 & 3.91 \\
\NAVEC\lgl & \VEC\lgl & 8  & 219 & 3.65 \\
\NASCA\lgl & \VEC\lgl & 5 & 219 & 2.28 \\
\bottomrule
\end{tabular}
% END Autogenerated
 \caption{What is the highest failure rate among popular argument types? For argument signatures whose frequency is in the 90th percentile}
 \label{tbl:fail-breakdown-2}
 \end{table}

  \begin{table}
% BEGIN Autogenerated
\scriptsize
\begin{tabular}{ll@{}rrr@{}}
\toprule
Passed & Arg Type & \#Args Fail & \#Args w Type & \% Failure\\
\midrule
\CLASS{{\tt matrix}} & \CLASS{{\tt timeSeries}} & 21 & 39 & 35.90 \\
\VEC\dbl & \CLASS{{\tt timeSeries}} & 21 & 39 & 35.90 \\
\CLASS{{\tt data.frame}} & \CLASS{{\tt timeSeries}} & 14 & 39 & 35.90 \\
\CLASS{{\tt dtplyr\_step\_first, dtplyr\_step}} & \CLASS{{\tt data.frame}} | Grouped Tibble | Tibble & 10 & 45 & 22.22 \\
\chr & \raw & 6 & 31 & 19.35 \\
% \CLASS{\texttt{data.frame}} & \CLASS{\texttt{data.frame}, \texttt{tbl}, \texttt{tbl\_df}} & 63 & 284 & 22.18\\
% \CLASS{\texttt{matrix}} & \CLASS{\texttt{array}} & 13 & 70 & 18.57\\
% \NAVEC\lgl & \lgl | \null & 9 & 49 & 18.37\\
% \VEC\chr & \LIST{\VEC\chr} & 20 & 109 & 18.35\\
% \VEC\clx & \clx & 7 & 42 & 16.67\\
\bottomrule
\end{tabular}
% END Autogenerated
 \caption{What is the highest failure rate among popular argument types? For argument signatures whose frequency is in the 80th percentile.
 *: Grouped Tibble is \CLASS{{\tt data.frame, grouped\_df, tbl, tbl\_df}} , and Tibble is \CLASS{{\tt data.frame, grouped\_df, tbl, tbl\_df}}.}
 \label{tbl:fail-breakdown-3}
 \end{table}

@@

 First, we were curious about how many \null-typed values flowed into non-nullable arguments, and we found that they accounted for \NUMNULLASSERTERRORS of contract assertion failures in \NUMFUNSWITHNULLASSERTERROR functions.
 We manually inspected some of the offending functions and observed three main patterns.

 First, we found that many of these errors occurred in arguments that have a {\tt NULL} default value.
 This would be the case when the programmer only tested the function by passing values to these null-by-default arguments, and clients of the package make use of the default.
 As an example, we observed a contract assertion failure when {\tt inner} and {\tt labels} were {\tt NULL} in the following function:
 \begin{lstlisting}
nlme::nfGroupedData <- function (formula, data = NULL, order.groups = TRUE, 
    FUN = function(x) max(x, na.rm = TRUE), outer = NULL, inner = NULL, 
    labels = NULL, units = NULL) {
  # ...
}
 \end{lstlisting}

@@

 Next, we were curious about how often vectors were passed to arguments expecting scalars.
 We found that \PERCDYNAMICVECERRORS of dynamic contract assertion failures were of this type, and these errors were present in \NUMFUNSWITHVECASSERTERROR functions.
 Besides an outright error, this kind of contract assertion failure might indicate that a function was not well-tested, in that it was only ever tested with unit-length vectors being passed to an argument which is intended to have a vector type.
 Further, these errors may reveal functions that were not designed with a vector-typed argument in mind, but can in fact handle vectors of values (in R, most functions that can accept scalars can also accept vectors).
 As an example, consider the function \code{BBmisc::strrepeat}, which takes a string and repeats it a specified number of times:
 \begin{lstlisting}
BBmisc::strrepeat <- function (x, n, sep = "") {
    paste0(rep.int(x, n), collapse = sep)
}

@@

Finally, we were curious to see what patterns of errors occurred in arguments expecting classes.
Overall we found that \PERCFAILASSCLASS of assertion failures were on arguments which were expecting a class in some way (\PERCFAILEDONECLASS of assertion failures were on monomorphic arguments expecting a class, the remainder on polymorphic arguments with at least one option being a class).

There are two broad divisions which account for most of the class-related contract assertion failures that are not outright errors.
First, we observe a class of errors related to ``subclasses'' being passed to arguments expecting a ``superclass''.
For instance, we observed \CLASS{{\tt data.frame}} values being passed to arguments expecting tibbles or data.tables (which can be readily converted).
Second, we observe a class of errors related more to coercion between simple data types and classes.
As an example, consider the aforementioned assertion failures in the \code{timeSeries} package, and as a further example we found many instances of \CLASS{{\tt matrix}}, \CLASS{{\tt data.frame}}, and vectors being passed to arguments expecting \CLASS{{\tt array}}, a generalization of matrices.

@@

 To further validate our inferred types, we repeated the experiment discussed in this section on end-user R code found on the Kaggle competition website.

 By-and-large, we saw no meaningful difference in the two data sets, with the contract assertion failure patterns being repeated from the reverse dependencies.
 Overall, we observed that \PERCFAILEDASSERTIONSKAGGLE of all contract assertions failed while running Kaggle code.
 If we remove assertion failures related to our simplifying assumption that function types will not have more than 20 arguments, that number drops to a mere \PERCFAILEDASSERTIONSKAGGLENOUNDEF.
 In all, \PROPFUNSWITHFAILEDASSERTKAGGLE functions had at least one contract failure.
 There were \NUMTOTALASSERTIONSKAGGLE assertions in total, on \NUMASSERTEDFUNCTIONSKAGGLE functions.

 \begin{table}[ht]
 \scriptsize
\centering
\begin{tabular}{llrrr}
\toprule
Passed & Arg Type & Occurrences & \% Total & Cumulative \%\\
\midrule
 \CLASS{{\tt data.table, data.frame}} & \CLASS{{\tt matrix}} & 20002 & 28.15 & 28.15 \\ 
 \CLASS{{\tt factor}} & \NAVEC\chr & 18344 & 25.81 & 53.96 \\ 
 \CLASS{{\tt factor}} & \VEC\chr  & 7519 & 10.58 & 64.54 \\ 
 \VEC\chr & \LIST{\VEC\int} $|$ ... $|$ \LIST{\CLASS{{\tt formula, quosure}}} & 5385 & 7.58 & 72.12 \\ 
 \CLASS{{\tt ixforeach, iter}} & \CLASS{{\tt dataframeiter, iter}} $|$ ... $|$ \CLASS{{\tt iforeach, iter}} & 4139 & 5.82 & 77.94 \\ 
\bottomrule
\end{tabular}
\caption{Top contract failures in Kaggle kernels}
\label{tbl:fail-breakdown-0-kaggle}
\end{table}

 \begin{table}
% BEGIN Autogenerated\
\small
\begin{tabular}{llrrr}
\toprule
Passed & Arg Type & \# Args Failed & \# Args with Type & \% Failure\\
\midrule
\VEC\chr & \chr & 12 & 304 & 0.04 \\ 
\CLASS{{\tt factor}} & \VEC\chr & 7 & 199 & 0.04 \\ 
\VEC\dbl & \VEC\chr & 6 & 199 & 0.03 \\ 
\NAVEC\chr & \VEC\chr & 4 & 199 & 0.02 \\ 
\int & \chr & 5 & 304 & 0.02 \\ 
\bottomrule
\end{tabular}
\caption{What is the highest failure rate among popular argument types in Kaggle? For argument signatures whose frequency is in the 90th percentile}
 \label{tbl:fail-breakdown-2-kaggle}
\end{table}

 To mirror our analysis of contract assertions on the reverse dependencies of our corpus, we show in Table~\ref{tbl:fail-breakdown-0-kaggle} the most frequently failing contract in Kaggle.
 While we don't see many overlapping entries per se, the table entries have similar explanations.
 For instance, values of type \CLASS{{\tt data.table, data.frame}} are often passed to arguments expecting a \CLASS{{\tt matrix}}.
 \CLASS{{\tt data.table, data.frame}} values are known as ``data tables'', and essentially serve the same purpose as data frames and tibbles.
 As it happens, many data tables can be coerced to matrices, and programmers will often interchange the two (particularly when data tables have numeric data), as is the case here.
 One function producing many of these errors is the following:
 \begin{lstlisting}
class::knn <- function (train, test, cl, k = 1, l = 0, prob = FALSE, use.all = TRUE)  {
    train <- as.matrix(train)
    test <- as.matrix(test)
    ...
}
 \end{lstlisting}
 We see that \code{class::knn} can take arguments as long as they can be passed to \code{as.matrix}, e.g., data tables, and will perform the coercion.
 As in similar cases discussed earlier, we did not encode this semantics in our type system as the process of coercion in R is ad hoc.
 On the topic of coercion, rows two and three of Table~ are interesting as they depict {\it factors} being passed to arguments expecting character vectors.
 \CLASS{{\tt factor}} typed values are known as factors in R, and they are stored as a vector of integer values corresponding to a set of character values, and their purpose is to allow for quickly deal with categorized data.
 Factors can be readily converted to characters when needed, as evidenced by these assertion failures.

 Another interesting entry in Table~\ref{tbl:fail-breakdown-0-kaggle} is the fifth row, where a \CLASS{{\tt ixforeach, iter}}  is being passed to an argument expecting a long union of classes, each of the form \CLASS{{\tt X, iter}}.
 This is likely an instance where the user defined their own class, {\tt ixforeach}, and wanted to use the \code{iterators} package (the user called \code{iterator::iter} with a \CLASS{{\tt ixforeach}}, and it gained the class tag {\tt iter} on return).
 As we mentioned, accounting for object-orientation in the type system is beyond the scope of this work, and such an inclusion would allow us to better type situations like this.

 Table~\ref{tbl:fail-breakdown-2-kaggle} mirrors Table~\ref{tbl:fail-breakdown-2} in showing contract violations on the most frequently occurring argument types.
 Here, our manual analysis has revealed similar failure patterns.
 In the case of vectors being passed to scalars, we find functions which can take vectors but were only tested with scalars (e.g., \code{stringr::str_to_upper} which converts a vector of characters to upper case, and \code{dplyr::anti_join}, which can join by a vector of column names but was only ever tested with a scalar).
 We also see possibly-NA character vectors being passed to NA-free character vectors.
 These assertion failures arise from a lack of testing: the offending functions are \code{str_trim}, \code{str_sub}, and \code{str_replace_all} from the \code{stringr} package.
 These functions are actually wrappers for other functions which have the correct argument type (\NAVEC\chr).

 \subsubsection{Discussion}

 Overall, the analysis discussed in this section has revealed two broad categories of contract assertion failures: those related to coercion, and those related to a class hierarchy.
 Our type system does not account for coercion as coercion in R is ad hoc at best, and it is implemented on a function-by-function basis, even in the core R packages.
 As for errors related to a class hierarchy, we aim to tackle this in future work, as designing a full fledged object-oriented type system for a language like R is outside of the scope of this work.

@@

### 2020-09-14 paper/tables/uni-typed-breakdown.tex
unlock 6 7
Add graph
new words in. let the edits commence

\begin{table}[H]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{lrrr}
\toprule
Types & Parameter \# & \% & Cumulative \%\\
\midrule
scalar & 28045 & 35.80 & 35.80\\
class & 18510 & 23.63 & 59.43\\
vector & 10296 & 13.14 & 72.58\\
\rowcolor{lightgray}  \textcolor{black}{...} & \textcolor{black}{6611} & \textcolor{black}{8.44} & \textcolor{black}{81.02}\\
\rowcolor{lightgray}  \textcolor{black}{any} & \textcolor{black}{5993} & \textcolor{black}{7.65} & \textcolor{black}{88.67}\\
\addlinespace
null & 4795 & 6.12 & 94.79\\
\rowcolor{lightgray}  \textcolor{black}{list} & \textcolor{black}{3229} & \textcolor{black}{4.12} & \textcolor{black}{98.91}\\
\textasciicircum{}vector & 461 & 0.59 & 99.50\\
environment & 390 & 0.50 & 100.00\\
\bottomrule
\end{tabular}}
\end{table}

>>>>

\begin{table}[H]
  \centering
  \resizebox{.5\linewidth}{!}{
    \begin{tabular}{lrrr}
      \toprule
      Types & Parameter \# & \% & Cumulative \%\\
      \midrule
      scalar & 29673 & 35.06 & 35.06\\
      class & 20308 & 23.99 & 59.05\\
      vector & 11529 & 13.62 & 72.67\\
      \rowcolor{lightgray}  \textcolor{black}{...} & \textcolor{black}{6611} & \textcolor{black}{7.81} & \textcolor{black}{80.48}\\
      \rowcolor{lightgray} \textcolor{black}{null} & \textcolor{black}{6094} & \textcolor{black}{7.20} & \textcolor{black}{87.68}\\
        \addlinespace
   \rowcolor{lightgray}  \textcolor{black}{any} & \textcolor{black}{6011} & \textcolor{black}{7.10} & \textcolor{black}{94.78}\\
      \rowcolor{lightgray}  \textcolor{black}{list} & \textcolor{black}{3400} & \textcolor{black}{4.02} & \textcolor{black}{98.79}\\
      \textasciicircum{}vector & 612 & 0.72 & 99.52\\
      environment & 408 & 0.48 & 100.00\\
      \bottomrule
    \end{tabular}}
  \caption{Singleton Type Categories}\label{fig:singlefreq}
\end{table}
